---
title: "Area description: Parish of St. Anselm, Hayes"
author: "Dave Lovell (Church Army's Research Unit)"
format: docx
editor: visual
echo: false
message: false
params:
  file: "2023-05-17_st-anselm-hayes.RDS"
  diocese: "London"
  nation: "England"
---

```{r libraries}
#| include: false
library(conflicted)
library(sf)
library(dplyr)
conflicted::conflicts_prefer(dplyr::filter)
library(ggplot2)
library(stringr)
library(knitr)
library(ggmap)
library(magrittr)
library(carutools)
library(here)
library(vroom)

plots_tables <- readRDS(here("plots-and-tables", params[["file"]]))

plots <- plots_tables[["plots"]]
tables <- plots_tables[["tables"]]
rm(plots_tables)
```

# Introduction

This report explores the statistical profiles of several areas within the parish of Hayes. These are:

-   Red Brick Estate

-   Townfield Estate

-   The parish of Hayes itself

These areas are plotted in Figure 1.

As well as being compared to one another, each of these areas will also be compared to the wider diocese (`r params[["diocese"]]`) and to `r params[["nation"]]` as a whole.

```{r areaMap}
#| fig-width: 6
#| cache: true
#| include: false

shapes <- read_sf(here("geospatial-reporting-data",
                       "st-anselm-hayes_2023-05-17",
                       "st-anselm-hayes_requested-shapes.geojson"))

shapes <- st_transform(shapes, 4326)

basemap <- get_map(
  st_bbox(shapes) |> 
    set_names(c("left", "bottom", "right", "top")),
  maptype = "satellite",
  zoom = 14)
```

```{r makeMapPlot}
#| include: false
areas_map <- 
  ggmap(basemap) + 
    geom_sf(data = shapes,
          aes(geometry = geometry, fill = name),
          size = 1, alpha = 0.2, colour = "black",
          inherit.aes = FALSE) + 
    ca_scale_fill_discrete(name = "Geography") +
    xlab(NULL) + ylab(NULL) + 
    theme(
      legend.position = "bottom",
      axis.line = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank()) + 
    labs(title = "Geographies analysed", caption = "(C) Google 2023")
```

```{r}
#| fig-width: 7
#| fig-cap: "Figure 1: Geographies described in this report"
areas_map
```

The domains of statistical data covered by this report are:

-   Ethnicity

-   Age

-   Religion

-   General health

-   Household composition

-   Deprivation

All the data in this report were calculated using the results of the Office for National Statistic's (ONS) 2021 Census of England and Wales. These data are available for output areas, which are the smallest statistical geography used by ONS. The statistics in this report were calculated by spatially joining the geographies of the respective estates and the parish with the output areas they intersect.

# Findings

## Overview of Findings

-   The parish and each of the estates are home to populations that are more ethnically diverse, more religious and younger than the national population.

-   Single occupant households account for more than a third of all households on Red Brick Estate

-   Both estates, but especially Townfield, are comprised of a large number of households in which not all occupants are part of a single nuclear family

-   Both estates and most of the parish are made up of neighbourhoods that are in the top 20% most deprived neighbourhoods in the UK

## Ethnicity

In each of the estates, white people account for 25% of the population or less. In the diocese and the nation, 82% of the population are white.

The ethnicity profiles of Townfield estate and the parish are similar, with around half the population identifying as Asian and 13% of the population identifying as Black. On Red Brick estate, only 35% of the population are Asian, and 30% of the population are black.

Figure 2 shows the variation in ethnicity across all geographies, and Table 1 provides a tabluation of this data.

```{r ethnicityPlot}
#| fig-width: 7
#| fig-cap: "Figure 2: Ethnicity across geographies"
plots[["ethnicity"]] + 
  labs(caption = "ONS census 2021. Church Army's Research Unit (DL) 2023")
```

```{r ethnicityTable}
#| tbl-cap: "Table 1: Ethnicity across geogrpahies"
tables[["ethnicity"]] |> 
  kable()
```

## Age

Both estates are home to a population that is younger than the national average: In Red Brick and Townfield, approximately 1 in 4 people are aged fourteen or below. Across the nation, this figure is closer to 1 in 6. Likewise, individuals over the age of 60 are underrepresented on both estates, and in the parish generally.

Figure 3 shows the breakdown of ages across all geographies, in fifteen-year intervals. This data is also tabulated in Table 2. Appendix 1 is a table that shows how ages vary across each geography in 5 year intervals.

```{r agePlot}
#| fig-width: 7
#| fig-cap: "Figure 3: Population ages in each geography"

plots[["age"]] + 
  labs(caption = "ONS census 2021. Church Army's Research Unit (DL) 2023")
```

```{r ageTable}
#| tbl-cap: "Table 2: Population ages in each geography"
tables[["age"]] |> 
  kable()
```

## General Health

Figure 4 shows the self-reported general health of residents in both estates is close to the national average, and only slightly worse than the average outcomes for the parish. A high percentage of the parish population are in 'good' or 'very good' health compared to the national average; this is partly explained by the youthfulness of the parish's population. This effect is less pronounced for the estates, despite their young populations. These data area also available in Table 3.

```{r generalHealthBar}
#| fig-width: 7
#| fig-cap: "Figure 4: General health in each geography"

plots[["general_health"]] + 
  labs(caption = "ONS census 2021. Church Army's Research Unit (DL) 2023")
```

```{r generalHealthTable}
#| tbl-cap: "Table 3: General health across geographies"
tables[["general_health"]] |> 
  kable()
```

## Religion

In religious terms, the greatest similarity between the Parish of St Anselm and the nation of England may be the abscense of a religious majority in either. Apart from this, their religious profiles differ starkly.

The parish is much less Christian and much less non-religious than the nation, and these deficits are made up by increases in the relative number of Sikhs, Hindus and Muslims. Both estates have more Muslims and fewer Hindus or Sikhs than the parish, and this effect is most pronounced on Red Brick estate, where almost 40% of residents are Muslim. Figure 5 and Table 4 show the religion data for all geographies in more detail.

```{r religionBar}
#| fig-width: 7
#| fig-cap: "Figure 5 - Religion across geographies"
plots[["religion"]]+ 
  labs(caption = "ONS census 2021. Church Army's Research Unit (DL) 2023")
```

```{r religionTable}
#| tbl-cap: "Table 4: Religion across geographies"
tables[["religion"]] |> 
  kable()
```

## Household composition

The subsequent sections describe the composition of households across the parish and both estates. A complete table of household composition data is available in index 2. Noteworthy headlines include:

-   Over one third of households on the Red Brick estate are single-occupant households

-   Childless couples are less prevalent on both estates that the national average

-   Households comprised of people from multiple families account for 21% of households in Townfield estate

### Lone occupants and single parents

Figure 6 shows the representation of lone-occupant and single-parent households across all geographies. Table 5 tabulates these figures.

Lone occupant households account for over a third (35%) of all households on Red Brick Estate. This is higher than the national average of 30%, despite the estate's considerably younger population. Single parents are more prevalent on both estates, and particularly those with dependent children.

```{r lonersBar}
#| fig-cap: "Figure 6: Lone-occupant and single-parent households"
#| fig-width: 7
plots[["household_composition"]][["loners"]] + 
  labs(caption = "ONS census 2021. Church Army's Research Unit (DL) 2023")
```

```{r lonersTable}
#| tbl-cap: "Table 5: Lone-occupant and single-parent households"
tables[["household_composition"]][["loners"]] |> 
  kable()
```

### Couples (married or cohabiting)

Figure 7 shows the representation of households consisting of married or co-habiting couples across all geographies. The prevalence of households comprised of a couple and their dependent children is slightly higher than the national average in both estates, where these households constitute approximately 1/5th of all households.

On both estates there are relatively few homes in which the sole occupants are a couple; these homes account for approximately one home in sixteen on each estate, as opposed to the national average of approximately one home in six. These figures are tabulated in more detail in Table 6.

```{r couplesPlots}
#| fig-cap: "Figure 7: Couples (married or cohabiting)"
#| fig-width: 7
plots[["household_composition"]][["couples"]] + 
  labs(caption = "ONS census 2021. Church Army's Research Unit (DL) 2023")
```

```{r copulesTable}
#| tbl-cap: "Table 6: Couples (married or cohabiting)"
tables[["household_composition"]][["couples"]] |> 
  kable()
```

### Other household structures

Figure 8 shows the representation of other households across all geographies. These households are comprised neither of couples (and their children), single-parents or lone-occupants.

The shortage of elderly families (all occupants over 66 years old) is unsurprising in a context where the elderly residents are so few. Particularly noteworthy, however, is the prevalence of household that do not fit neatly into the categories proposed by the census. These households may consist of extended families living together, of families with lodgers or of flat-shares and student accommodation. Such households are especially prevalent on Townfield estate, where they comprise 21% of all households. Table 7 shows these data in more detail.

```{r otherHouseholdsPlot}
#| fig-cap: "Figure 8: Other households"
#| fig-width: 7
plots[["household_composition"]][["others"]] + 
  labs(caption = "ONS census 2021. Church Army's Research Unit (DL) 2023")
```

```{r otherHouseholdsTable}
#| tbl-cap: "Table 7: Other household types"
tables[["household_composition"]][["others"]] |> 
  kable()
```

## Deprivation

Overall deprivation across the parish and in both estates is high. Most of the parish population live in neighbourhoods that are classified as being in the 20% most neighborhoods of the UK. Figure 9 shows the intersection of regional deprivation with the parish and the estates.

Deprivation scores are from the 2019 Index of Multiple Deprivation (IMD), which compares and ranks lower-layer super output areas across England and Wales on the basis of multiple weighted indices of deprivation, including income, employment, education and health. For more information on the 2019 IMD, see [this infographic](https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/833959/IoD2019_Infographic.pdf).

```{r loadLsoa}
#| include: false
lsoa <- read_sf(here("civic-areas",
                     "2011_lsoa-boundaries_england-wales_generalised-clipped"))
lsoa <-
  select(lsoa, lsoa_2011_code = LSOA11CD) |> 
  filter(str_starts(lsoa_2011_code, "E"))

lsoa <- st_transform(lsoa, st_crs(shapes))
```

```{r loadIMDData}
#| include: false
imd <- vroom(here("small-area-stats", "imd_2019.csv"),
             show_col_types = FALSE) |> 
  select(lsoa_2011_code = `LSOAcode(2011)`,
         imd_decile = `IndexofMultipleDeprivation(IMD)Decile`,
         imd_rank = `IndexofMultipleDeprivation(IMD) Rank`)

imd <- filter(imd, str_starts(lsoa_2011_code, "E"))

imd <- mutate(imd, imd_rank = rank(imd_rank))

imd <- 
  mutate(imd, 
         imd_percentile =  ceiling(100 * imd_rank / max(imd_rank)),
         imd_twentieth = ceiling(imd_percentile/5))

imd <- 
  mutate(imd,
         custom_label = 
           case_when(
             imd_twentieth == 1 ~ "5% most dep.",
             imd_twentieth == 2 ~ "10% most dep.",
             imd_twentieth == 3 ~ "15% most dep.",
             imd_twentieth == 4 ~ "20% most dep.",
             TRUE               ~ "other") |>
           
           ordered(levels = c("5% most dep.", "10% most dep.",
                              "15% most dep.", "20% most dep.",
                              "other")))
```

```{r findIntersectingOa}
#| include: false
intersecting_lsoa <-
  st_intersects(lsoa, st_union(shapes), sparse = FALSE) |> 
  as.vector()

lsoa <- 
  filter(lsoa, intersecting_lsoa) |> 
  left_join(imd, by = "lsoa_2011_code")

imd_map <-
  ggmap(basemap) + 
  geom_sf(data = lsoa,
          aes(fill = custom_label),
          colour = "black", linewidth = 0.8,
          inherit.aes = FALSE,
          alpha = 0.7) +
  # geom_sf_text(data = lsoa,
  #              aes(label = imd_percentile),
  #              colour = "gray30",
  #              inherit.aes = FALSE) + 
  scale_fill_manual(name = "IMD Rank",
                    values =
                      colorRampPalette(c(ca_dark_teal(), "white"))(5)[1:4] |>
                      c(NA),
                    drop = FALSE,
                    na.value = NA) + 
    geom_sf(data = shapes,
          aes(colour = name),
          fill = NA, linewidth = 1.2,
          inherit.aes = FALSE) +
  scale_colour_manual(values = ca_orange(), name = "Parish") +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  ggtitle("Deprivation in both parishes") + 
  labs(caption = "IMD 2019. Church Army's Research Unit (DL) 2023")
```

```{r printMap}
#| fig-width: 7 
#| fig-cap: "Deprivation in LLSO that intersect the parish & estates"

imd_map
```

# Appendices

## Appendix 1: 5-year age table

```{r detailedAgeTale}
tables[["age_five_year"]] |> 
  kable()
```

## Appendix 2: All household composition data

```{r householdCompDataAll}
tables[["household_composition"]][["all"]] |> 
  kable()
```
