---
title: "Splitting St Paul's Parish in Half"
author: "Dave & Shiobhan"
format: html
editor: visual
---

## Introduction 

## Libraries

```{r libraries}
library(fs)
library(here)

library(sf)
library(ggplot2)
library(dplyr)
library(vroom)
```

## Load Data

```{r load_data}
# parish <- read_sf(here("parish-shapes", "coe-parishes.geojson"), )
llsoa <- read_sf(here("civic-areas", "2011_lsoa-boundaries_england-and-wales_generalised-clipped"))
oa <- read_sf(here("2021_oa-boundaries_england-wales_generalised_clipped"))

imd <- vroom(here("small-area-stats", "imd_2019.csv"))
```

## Find intersecting LLSOA 

### Reproject parish shape

We need to reproject `parish` so that it shares the same projection as `llsoa`. This will enable us to find those LLSOA that intersect the parish.

```{r reproject_parish}

```

### Find intersecting LLSOA

Use `st_intersects()` to determine which LLSOA intersect the parish

```{r find_intersecting_llsoa}

```

## Join IMD data to LLSOA

`left_join()` IMD data to `llsoa` by LLSOA code
