---
title: "Jesus Shaped People: St Paul's, Bristol"
author: "Dave"
format: html
editor: visual
params: 
  nation: "england"
  diocese: "bristol"
---

## Housekeeping

### Libraries

```{r libs}
#| output: false
library(conflicted)

library(fs)
library(here)

library(vroom)
library(stringr)
library(purrr)
library(sf)
library(dplyr)
conflicted::conflicts_prefer(dplyr::filter)
library(ggplot2)
library(lobstr)
library(tidyr)
library(scales)
library(forcats)
library(carutools)
```

### Data, files and paths

### Get shapes

#### Get shapes for requested areas

Read in both `.kml` files, then bind all three shapes into one `sf` object.

```{r get_praish_shapes}
requested_area <-
  read_sf(
    here("requested-areas", "2023-09-26_st-pauls-bristol-split.geojson")
    )

```

#### Get shapes for small areas

```{r getSmallAreaShapes}
oa <-
  read_sf(here("civic-areas", "2021_oa-boundaries_england-wales_generalised_clipped")) |> 
  select(oa_2021_code = OA21CD)
```

#### Reproject requested shapes

```{r reproject}
requested_area <- st_transform(requested_area, st_crs(oa))
```

### Get statistics

#### Get diocesan stats

Get stats for `r params[["diocese"]]`

```{r getDiocesanStats}

stat_names <- c("age-five-year", "ethnicity", "general-health", "household-composition", "religion")

diocesan_stats <- 
  map(stat_names,
      function(stat_name){
        
        file_name <- str_c(params[["nation"]], "_", stat_name, ".csv")
        file_path <- here("diocesan-stats", file_name)
        
        vroom(file_path, show_col_types = FALSE) |> 
          filter(diocese == str_to_title(params[["diocese"]])) |> 
          select(-diocese)
      })

names(diocesan_stats) <- str_replace_all(stat_names, "-", "_")
```

#### Get national stats

```{r getNationalStats}
oa_stats <- 
  map(stat_names,
      function(stat_name){
        
        file_name <- str_c("england-wales_", stat_name, "_oa_2021.RDS")
        file_path <- here("small-area-stats", file_name)
        
        readRDS(file_path)
      })

names(oa_stats) <- str_replace_all(stat_names, "-", "_")

## NB: about 5 'general-health' stats contain NA values for the
## whole of England. This stops those from making this crash.
oa_stats$general_health |>
  filter(if_any(where(is.numeric), is.na))

national_stats <- 
  map(oa_stats, 
      \(x){
        filter(x, if_all(where(is.numeric), negate(is.na))) |> 
        summarise(across(where(is.numeric), sum))
        })
```

## Format Data

### Match OA to LLSOA

LLSOA are comprised of OA. Lets:

-   Get the union of the parish

-   Find all OA that intersect the union

-   Get respective unions of the North and South of the parish

-   Match OA to N/S on the basis of contained centroids

### Filter intersecting civic areas

We only need to deal with output areas that intersect the shapes we're analysing.

```{r filterOA}
request_union <- st_union(requested_area)

intersecting_oa <-
  st_intersects(oa, request_union, sparse = FALSE) |> 
  as.vector()

oa <- filter(oa, intersecting_oa)
oa <- mutate(oa, oa_area = as.vector(st_area(geometry)))
nrow(oa)

ggplot(oa) + 
  
  geom_sf(fill = ca_cyan(), alpha = 0.5, colour = "black") + 
  
  geom_sf(data = requested_area,
          aes(fill = south),
          colour = ca_purple(),
          alpha = 0.7)
```

Great, that looks like a centroid-belonging join will work.

### Get intersections and centroids

We can use the centroids of intersected areas to determine whether an oa belongs to the North or the South

```{r get_centroids}
oa_llsoa_intersection <- st_intersection(oa, requested_area)

oa_intersection_centroids <- st_centroid(oa_llsoa_intersection)

ggplot(requested_area) + 
  geom_sf(fill = ca_cyan(), colour = "black", alpha = 0.6) + 
  geom_sf(data = oa_intersection_centroids, colour = ca_gold())
```

### Make N/S unions

```{r n_s_unions}
north_south <- 
  group_by(requested_area, south) |> 
  summarise()
```

### Find respective distances of each OA from N/S

```{r find_distance_to_n_s}
which_ordinal <- if_else(north_south[["south"]], "south", "north")
n_s_distances <- st_distance(oa_intersection_centroids, north_south)
colnames(n_s_distances) <- which_ordinal

head(n_s_distances)

north_or_south <-
  apply(n_s_distances, 1, \(x) which_ordinal[x == min(x)]) |> 
  unlist()

oa_intersection_centroids <-
  mutate(oa_intersection_centroids,
         north_or_south = north_or_south)
```

### Assign ordinal category to each OA

```{r assign_oa_ordinals}
oa_intersection_centroids <- 
  as_tibble(oa_intersection_centroids) |>
  select(oa_2021_code, north_or_south)

oa <-
  left_join(oa, oa_intersection_centroids, by = "oa_2021_code") |> 
  mutate(north_or_south = factor(north_or_south))

ggplot(oa, aes(fill = north_or_south)) + 
  geom_sf()

st_intersection(oa, request_union) |> 
  ggplot(aes(fill = north_or_south)) + 
  geom_sf()


```

## Get OA intersections

```{r get_oa_intersections}
oa_intersections <- st_intersection(oa, request_union)

oa_intersections <- mutate(oa_intersections, area = as.vector(st_area(geometry)))

oa_intersections <- mutate(oa_intersections, oa_in_area = as.vector(area/oa_area), .keep = "unused")
```
### Save shape data

```{r saveShapeData}
respective_unions <- 
  requested_area |> 
  group_by(south) |> 
  summarise()

grd <- dir_create(here("geospatial-reporting-data"))
hayes <- dir_create(grd, "st-phillip-upper-stratton_2023-09-11")

## using this because write_sf appears to be borked
st_write(oa_intersections, here(hayes, "st-pauls-bristol-split_oa.geojson"),
         delete_dsn = TRUE)

st_write(respective_unions, here(hayes, "st-pauls-bristol-split.geojson"),
         delete_dsn = TRUE)
```
## Calculate statistics

### Format requested shapes as list

```{r calculateStats}
requested_area <- group_by(oa_intersections, north_or_south)

shape_keys <-
  group_keys(requested_area) |> 
  unlist()

request_list <- 
  group_split(requested_area) |> 
  set_names(shape_keys)

rm(shape_keys)

lobstr::tree(request_list, max_depth = 1)
```

### Map across shapes & stats

```{r mapShapesStats}
requested_area_stats <- 
  map(request_list,
    function(x){
      map(oa_stats, 
          function(y){
            left_join(x, y, by = "oa_2021_code") |> 
              as_tibble() |> 
              select(-geometry) |> 
              mutate(across(where(is.numeric), \(a) a * oa_in_area)) |> 
              select(-oa_in_area) |> 
              summarise(across(where(is.numeric), sum))
          })
    })

lobstr::tree(requested_area_stats, max_depth = 2)
```

### Combine all areas

```{r combineAllAreas}
all_areas <- c(requested_area_stats,
               list(diocese = diocesan_stats),
               list(nation  = national_stats))

lobstr::tree(all_areas, max_depth = 2)

cool_crs <- 

all_areas <- 
  imap(all_areas,
     \(area_list, area_name){
       map(area_list,
           \(x){
             neat_area_name <- str_to_lower(area_name)
             
             mutate(x, area = neat_area_name) |> 
               relocate(area)
             }) })

### Parrallel-map across each list item, reducing their elements into a single tibble
all_areas <- pmap(all_areas, \(...) reduce(list(...), rbind))

all_areas <- map(all_areas,
                 \(x){
                   mutate(x, area =
                            ordered(area) |> 
                            fct_inorder())
                 })
```

### Transform data

#### Make percentages

```{r makePercentages}
all_areas_abs <- all_areas
all_areas_rel <-
  map(all_areas,
      \(x){
        rowwise(x) |> 
          mutate(across(
            where(is.numeric),
            \(a) a / sum(c_across(where(is.numeric)))
            ))
      })

all_areas[[1]][1:4]
rowSums(all_areas[[1]][,-1])
```

#### Pivot Longer

```{r pivotLonger}
all_areas_rel <- map(all_areas_rel,
                 \(x){
                   pivot_longer(x,
                                cols = -area, 
                                names_to = "statistic", 
                                values_to = "proportion")
                   })

all_areas_abs <- map(all_areas_abs,
                 \(x){
                   pivot_longer(x,
                                cols = -area, 
                                names_to = "statistic", 
                                values_to = "proportion")
                   })
```

## Save data

```{r saveData}
save_dir <- dir_create(here("comparative-tables"))

saveRDS(all_areas_rel, here(save_dir, "2023-09-27_st-pauls-bristol-split_proportions.RDS"))
saveRDS(all_areas_abs, here(save_dir, "2023-09-27_st-pauls-bristol-split_population.RDS"))
```
